Ground state energy of quantum dots using the coupled cluster method

### 2018-04-17 project_2/article/article.tex
Start on two-body normal ordering

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        v(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + W
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $W$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or de-exciting the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
        \end{align}
        to find $t_{ij}^{ab}$.

@@

\appendix
\section{The normal ordered Hamiltonian}
    When constructing the normal ordered Hamiltonian we use Wick's theorem to
    write the one-body, $h$, and the two-body, $W$, operators onto a normal
    ordered form. Specifically we define the normal ordered form in terms of
    Fermi vacuum. That is, an operator on normal ordered form destroys the
    reference Slater determinant.
    \begin{align}
        h &= \sum_{pq}h^p_q\acr{p}\ade{q}
        = \sum_{pq}h^p_q\para{
            \{\acr{p}\ade{q}\}
            + \{
                \wick{\c a_{p}^{\dagger} \c a_{q}}
            \}
        }
        \\
        &= \sum_{pq}h^p_q\{
            \acr{p}\ade{q}
        \}
        + \sum_{pq}h^p_q\delta_{p \in i}\delta_{pq}
        \\
        &= h_N + \sum_{i}h^i_i,
    \end{align}
    where we have used $\delta_{p \in i}$ to mean that $p$ must be an occupied
    index.
    \cite{crawford2007introduction}
    % TODO: Fill in the rest of this expansion

>>>>

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + W
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $W$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or de-exciting the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
        \end{align}
        to find an expression that can be used to solve for $t_{ij}^{ab}$.

@@

\appendix
\section{The normal ordered Hamiltonian}
    When constructing the normal ordered Hamiltonian we use Wick's theorem to
    write the one-body, $h$, and the two-body, $W$, operators onto a normal
    ordered form. Specifically we define the normal ordered form in terms of the
    \emph{Fermi vacuum}\footnote{Fermi vacuum defines the reference state, i.e.,
    $\kslat$, as the vacuum.}. That is, an operator on normal ordered form
    destroys the reference Slater determinant.

    We start by writing the one-body operator, $h$, to its normal-ordered form.
    \begin{align}
        h &= \sum_{pq}h^p_q\acr{p}\ade{q}
        = \sum_{pq}h^p_q\para{
            \{\acr{p}\ade{q}\}
            + \{
                \wick{\c a_{p}^{\dagger} \c a_{q}}
            \}
        }
        \\
        &= \sum_{pq}h^p_q\{
            \acr{p}\ade{q}
        \}
        + \sum_{pq}h^p_q\delta_{p \in i}\delta_{pq}
        \\
        &= h_N + \sum_{i}h^i_i,
    \end{align}
    where we have used $\delta_{p \in i}$ to mean that $p$ must be an occupied
    index. Doing the same for the two-body operator is a slightly more tedious
    endeavor. For brevity we will only write out the operator strings and only
    keep the non-zero contributions.
    \begin{align}
        \acr{p}\acr{q}\ade{s}\ade{r}
    \end{align}
    \cite{crawford2007introduction}
